---
output: github_document
format: gfm
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rinfa

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/rinfa)](https://CRAN.R-project.org/package=rinfa)
[![R-CMD-check](https://github.com/simonpcouch/rinfa/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/simonpcouch/rinfa/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of rinfa is to provide Rust bindings for parsnip model specifications.

## Installation

You can install the development version of rinfa from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("simonpcouch/rinfa")
```

## Example

```{r, echo = FALSE, message = FALSE}
# this code just logs the state of the parsnip model environment
# befor eloading rinfa so it can be used in the "Available implementations" sec
library(parsnip)

parsnip_models <- get_from_env("models") %>%
  setNames(., .) %>%
  purrr::map_dfr(get_from_env, .id = "model")

```

```{r}
x <- matrix(rnorm(3e7), ncol = 3, dimnames = list(NULL, paste0("X", 1:3)))
y <- rnorm(1e7)

dat <- cbind(as.data.frame(x), y)

system.time(
  lm(y ~ ., dat)
)

system.time(
  # lm()'s speedy friend, from base R
  .lm.fit(x, y)
)

library(rinfa)

system.time({
  .linfa_linear_reg(x, y)
})
```

To use rinfa with tidymodels, set the modeling engine to `"linfa"`:

```{r}
# using the formula interface:
linfa_fit <- fit(linear_reg(engine = "linfa"), y ~ ., dat)

# using the (more performant, in this case) XY interface:
linfa_fit_xy <- fit_xy(linear_reg(engine = "linfa"), x = x, y = y)
```

## Available implementations

The rinfa package provides an additional engine `"linfa"` for the models in the following table:

```{r, echo = FALSE, message = FALSE}
rinfa_models <- get_from_env("models") %>%
  setNames(., .) %>%
  purrr::map_dfr(get_from_env, .id = "model")

dplyr::anti_join(
  rinfa_models, parsnip_models, 
  by = c("model", "engine", "mode")
) %>%
  knitr::kable()
```

To read more about the design of rinfa, see [`R/README.md`](https://github.com/simonpcouch/rinfa/tree/main/R/README.md).
